MCC Models

### 2020-03-10 papers/full_paper/manuscript.Rmd
update paper after revision

We start by replicating previous findings with adults showing that listeners make pragmatic inferences based on non-linguistic properties of utterances in isolation (experiment 1).  Then we show that adults make inferences based on common ground information (experiment 2A and 2B). We use data from these experiments as parameters to generate a priori predictions from RSA models about how utterance and common ground information should be integrated. We consider three models that make different assumptions about the integration process: In the pragmatics model, the two information sources are integrated with one another; according to the flat prior model, participants focus only on the utterance information and in the prior only model, only common ground information is considered. We compare predictions from these models to new empirical data from experiments in which utterance and common ground information are manipulated simultaneously (Experiment 3 and 4).

@@

Adult participants were recruited via Amazon Mechanical Turk (MTurk) and received payment equivalent to an hourly wage of ~ \$9. Each participant contributed data to only one experiment. Experiment 1 and each manipulation of experiment 2 had *N* = 40 participants. Sample size in experiment 3 was *N* = `r length(unique(adult_ex3_data$id))`. *N* = `r length(unique(adult_ex4_prior_data$id))` participated in the experiments to measure the strong, medium and weak preference and novelty manipulations that went into experiment 4. Finally, experiment 4 had *N* = `r length(unique(adult_ex4_data$id))` participants. Sample sizes in all adult experiments were chosen to yield at least 120 datapoints per cell. All studies were approved by the Stanford Institutional Review Board (protocol no. 19960).

@@

Experiments 1 and 2 confirmed that adults make pragmatic inferences based on information provided by the utterance as well as by common ground and provided quantitative estimates of the strength of these inferences for use in our model. We modeled the integration of utterance informativity and common ground as a process of socially-guided probabilistic inference, using the results of experiments 1 and 2 to inform key parameters of a computational model. The Rational Speech Act (RSA) model architecture introduced by [@frank2012predicting] encodes conversational reasoning through the perspective of a listener (“he” pronoun) who is trying to decide on the intended meaning of the utterance he heard from the speaker (“she” pronoun). The basic idea is that the listener combines his uncertainty about the speaker’s intended meaning - a prior distribution over referents P(r) - with his generative model of how the utterance was produced: a speaker trying to convey information to him. To adapt this model to the word learning context, we enrich this basic architecture with a mechanism for expressing uncertainty about the meanings of words (lexical uncertainty) - a prior distribution over lexica P(L) [@bergen2016pragmatic].

@@

ex3_pragm_model_noise_param <- readRDS("../../stats/saves/ex3_pragm_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))%>%
  round(digits = 2)
```

In experiment 3, we combined the procedures of experiment 1 and 2A or 2B. The test setup was identical to experiment 1, however, before making a request, the speaker interacted with the objects so that some of them were preferred by or new to them (Fig \ref{fig:fig1}c). We discuss and visualize the results as the proportion with which participants chose the more informative object (i.e., the object that would be the more informative referent when only utterance information is considered).

@@

Participants distinguished between congruent and incongruent trials when the speaker remained the same, as evidenced by the fit of a generalized linear mixed effects model (model term: `alignment x speaker`; $\beta$ = `r lm_ex3_r%>%filter(term == "speakersame_speaker:alignmentincongruent")%>%pull(estimate)`, se = `r lm_ex3_r%>%filter(term == "speakersame_speaker:alignmentincongruent")%>%pull(std.error)`, *p* `r lm_ex3_r%>%filter(term == "speakersame_speaker:alignmentincongruent")%>%pull(p.value)`).

@@

Participants’ average responses were highly correlated with the model’s predictions in each condition (Fig \ref{fig:fig3}b). To test whether participants in fact balanced both information sources, we compared the pragmatics model to two alternative models: the *flat prior model*, which ignores common ground information and the *prior only model*, which ignores utterance information. Model fit was considerably better for the pragmatics model compared to the flat prior model (Bayes Factor (BF) = `r format(ex3_bf%>%pull(pragmatic_noise_flat_prior_noise), digits = 2)`) or the prior only model (BF = `r format(ex3_bf%>%pull(pragmatic_noise_prior_only_noise), digits = 2)`), suggesting that participants considered and integrated both sources of information. The estimated proportion of random responses according to the pragmatics model was `r ex3_pragm_model_noise_param%>%pull(mean)` (95% Highest Density Interval (HDI): `r ex3_pragm_model_noise_param%>%pull(ci_lower)` - `r ex3_pragm_model_noise_param%>%pull(ci_upper)`). This value was substantially lower for the pragmatics model compared to the alternative models (see Supplementary Material available online), lending additional support to the conclusion that the pragmatics model better captured the behavioral data. Rather than explaining systematic structure in the data, the alternative models achieved their best fit only by assuming a very high level of noise.

@@

```{r model comparison ex4}
ex4_bf <- bind_rows(
  readRDS("../../stats/saves/ex4_prior_only_model_noise_loglike.rds") %>% mutate(model = "prior_only_noise"),
  readRDS("../../stats/saves/ex4_pragm_model_noise_loglike.rds") %>% mutate(model = "pragmatic_noise"),
  readRDS("../../stats/saves/ex4_flat_prior_model_noise_loglike.rds") %>% mutate(model = "flat_prior_noise")
)%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))%>%
  spread(model, logP) %>%
  mutate("pragmatic_noise_prior_only_noise" = exp(pragmatic_noise - prior_only_noise),
         "pragmatic_noise_flat_prior_noise" = exp(pragmatic_noise - flat_prior_noise),
         "prior_only_noise_flat_prior_noise" = exp(prior_only_noise - flat_prior_noise)) %>%
  select(-pragmatic_noise, -flat_prior_noise, -prior_only_noise)
ex4_pragm_model_noise_param <- readRDS("../../stats/saves/ex4_pragm_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))%>%
  round(digits = 2)
```

To test if our model makes accurate predictions for different combinations, we first replicated and then extended the results of experiment 3 to a broader range of experimental conditions. Specifically, we manipulated the strength of the common ground information (3 levels - strong, medium and weak - for preference and 2 levels - strong and medium - for novelty) by changing the way the speaker interacted with the objects prior to the request. The procedural details and statistical analysis for these these manipulations are described in detail in the Supplementary Material available online. For experiment 4, we paired each level of prior strength manipulation with the informativeness inference in the same way as in experiment 3. This resulted in a total of 20 conditions, for which we generated a priori model predictions in the same way as in experiment 3. The strong prior manipulation in experiment 4 was a direct replication of experiment 3 (see Fig \ref{fig:fig3}c). Each participant was randomly assigned to a common ground manipulation and a level of prior strength and completed eight trials in total, two in each unique condition in that combination.

@@

GLMM results for experiment 4 can be found in the Supplementary Material available online. Here we focus on the analysis based on the probabilistic models. Model predictions from the pragmatics model were again highly correlated with the average response in each condition (see Fig \ref{fig:fig3}d). We evaluated model fit for the same models as in experiment 3 and found again that the pragmatics model fit the data much better compared to the flat prior (BF = `r format(ex4_bf%>%pull(pragmatic_noise_flat_prior_noise), digits = 2)`) or the prior only model (BF = `r format(ex4_bf%>%pull(pragmatic_noise_prior_only_noise), digits = 2)`). The inferred level of noise based on the data for the pragmatics model was `r ex4_pragm_model_noise_param%>%pull(mean)` (95% HDI: `r ex4_pragm_model_noise_param%>%pull(ci_lower)` - `r ex4_pragm_model_noise_param%>%pull(ci_upper)`), which was similar to experiment 3 and again lower compared to the alternative models (see Supplementary Material available online).

@@

The previous section showed that competent language users flexibly integrate information during pragmatic word learning. Do children make use of multiple information sources during word learning as well? When does this integration emerge developmentally? While many verbal theories of language learning imply that this integration takes place, the actual process has neither been described nor tested in detail. Here we provide an explanation in the form of our pragmatics model and test if it is able to capture children’s word learning. Embedded in the assumptions of the model is the idea that developmental change is change in the strength of the individual inferences, leading to a change in the strength of the integrated inference. As a starting point, our model assumes developmental continuity in the integration process itself, though this assumption could be called into question by a poor model fit. The study for children followed the same general pattern as the one for adults. We generated model predictions for how information should be integrated by first measuring children’s ability to use utterance (informativeness) and common ground (preference) information in isolation when making pragmatic inferences. We then adapted our model to study developmental change: We sampled children continuously between 3.0 and 5.0 years of age – a time in which children have been found to make the kind of pragmatic inferences we studied here [@bohn2019pervasive; @frank2014inferring] - and generated model predictions for the average developmental trajectory in each condition.

@@

In experiment 6, we assessed whether children use common ground information to identify the referent of a novel word. We tested children with the novelty as well as the preference manipulation but found little evidence that children distinguished between requests made by the same speaker or a different speaker in the case of novelty. Since our focus was on how children selectively integrate the two sources of information, we therefore dropped this manipulation and focused on preference for the remainder of the study. The procedure for children was identical to the preference manipulation for adults. Children received eight trials, four with the same and four with a different animal returning.

@@

In experiment 7, we combined the procedures of experiment 5 and 6 and collected new data from children between 3.0 and 5.0 years of age in each of the four conditions (Fig \ref{fig:fig1}c). We again inserted the preference manipulation into the setup of experiment 5. After greeting the child, the animal turned to one of the trees, pointed to an object (object was temporarily enlarged and moved closer to the animal) and expressed liking or disliking. Then the animal turned to the other tree and expressed the other attitude for the other kind of object. Next, the animal disappeared and either the same or a different animal returned. The rest of the trial was identical to the label and request phase of experiment 5. Children received eight trials, two per condition (same/different speaker x congruent/incongruent) in a randomized order.

@@

Children’s propensity to differentiate between congruent and incongruent trials for the same or a different speaker increased with age (model term: `age x alignment x speaker`; $\beta$ = `r child_lm_ex3_r%>%filter(term == "age_num:speakersame_speaker:alignmentincongruent")%>%pull(estimate)`, se = `r child_lm_ex3_r%>%filter(term == "age_num:speakersame_speaker:alignmentincongruent")%>%pull(std.error)`, *p* `r child_lm_ex3_r%>%filter(term == "age_num:speakersame_speaker:alignmentincongruent")%>%pull(p.value)`).

@@

child_ex3_pragm_model_noise_param <- readRDS("../../stats/saves/child_ex3_pragm_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))
```

Analyses comparing the model predictions from the probabilistic model to the data suggest that children flexibly integrate both common ground and informativity information. Furthermore, this integration process is accurately captured by the pragmatics model at least for four-year-olds. For the correlational analysis, we binned model predictions and data by year. There was a substantial correlation between the predicted and measured average response for four-year-olds, but less so for three-year-olds (Fig \ref{fig:fig5}b). One of the reasons for the latter was the low variation between conditions. For the model comparison, we treated age continuously. As with adults, we found a much better model fit for the pragmatics model compared to the flat prior (BF = `r format(child_ex3_bf%>%filter (parameter=="noise")%>%pull(pragmatic_flat_prior), digits = 2)`) or the prior only model (BF = `r format(child_ex3_bf%>%filter (parameter=="noise")%>%pull(pragmatic_prior_only), digits = 2)`). The inferred level of noise based on the data for the pragmatics model was `r child_ex3_pragm_model_noise_param%>%pull(mean)` (95% HDI: `r child_ex3_pragm_model_noise_param%>%pull(ci_lower)` - `r child_ex3_pragm_model_noise_param%>%pull(ci_upper)`), which was lower compared to the alternative models considered but numerically higher than that of adults (see Supplementary Material available online).

The high level of inferred noise moved the model predictions for children in all conditions close to chance level. We therefore compared two additional sets of models with different parameterizations that emphasized differences between conditions in the model predictions more (see Supplementary Material available online and Fig \ref{fig:fig5}a). This analysis was not pre-registered. Parameter free models did not include a noise parameter and developmental noise models allowed the noise parameter to change with age. In each case, the pragmatics model provided a better fit compared to the alternative models (flat prior: parameter free BF = `r format(child_ex3_bf%>%filter (parameter=="parameter free")%>%pull(pragmatic_flat_prior), digits = 2)`, developmental noise BF = `r format(child_ex3_bf%>%filter (parameter=="developmental noise")%>%pull(pragmatic_flat_prior), digits = 2)`; prior only: parameter free BF = `r format(child_ex3_bf%>%filter (parameter=="parameter free")%>%pull(pragmatic_prior_only), digits = 2)`, developmental noise BF = `r format(child_ex3_bf%>%filter (parameter=="developmental noise")%>%pull(pragmatic_prior_only), digits = 2)`). The developmental noise parameter for the pragmatics model decreased with age, suggesting that older children behaved more in line with model predictions compared to younger children (see Fig. S13 in Supplementary Material available online).

@@

Integrating multiple sources of information is an integral part of human communication [@tanenhaus1995integration]. To infer the intended meaning of an utterance, listeners must combine their knowledge of communicative conventions (semantics and syntax) with social expectations about their interlocutor. This integration is especially vital in early language learning, and the different varieties of pragmatic information are among the most important sources [@bohn2019pervasive]. But how are different cues integrated during word learning? Here we used a Bayesian pragmatics model to formalize this integration process. We studied how utterance-level about informative communication are integrated with common ground information. Adults’ and children’s learning was best predicted by a model in which both sources of information traded-off flexibly. Alternative models that considered only one source of information made substantially worse predictions.

All of the models we compared here integrated some explicit structure, rather than (for example) simply weighing expectations by some ratio. We made this decision because we wanted to make predictions within a framework in which the models were models of the task, rather than simply models of the data. That is, inferences are not computed separately by the modeler and specified as inputs to a regression model, but instead are the results of an integrated process that operates over a (schematic) representation of the experimental stimuli. Further, our models are variants derived from the broader RSA framework, which has been integrated into larger systems for language learning in context [@wang2016learning].

We conceptualized developmental change as age related changes in the propensity to make the individual inferences. That is, while the degree to which listeners expect speakers to be informative or follow common ground changes with age, the process by which expectations are integrated remains the same. However, other developmental models are also worth exploring in future work; one possible candidate would be a model in which the integration process itself changes with age. Our model did not successfully describe three-year-olds’ inferences; thus, it is possible that they were not able to integrate information sources. But our findings are also consistent with a simpler explanation, namely that the overall weaker responses we observed in the independent measurement experiments (experiments 5 and 6), combined with some noise in responding, led the younger children to appear relatively random in their responses.

@@

>>>>

We start by replicating previous findings with adults showing that listeners make pragmatic inferences based on non-linguistic properties of utterances in isolation (experiment 1). Then we show that adults make inferences based on common ground information (experiment 2A and 2B). We use data from these experiments as parameters to generate a priori predictions from RSA models about how utterance and common ground information should be integrated. We consider three models that make different assumptions about the integration process: In the *integration model*, the two information sources are integrated with one another. The other two models are lesion models that assume that participants focus on one type of information and disregard the other whenever they are presented together. According to the *no common ground* model, participants focus only on the utterance information and in the *no informativeness* model, only common ground information is considered. We compare predictions from these models to new empirical data from experiments in which utterance and common ground information are manipulated simultaneously (Experiment 3 and 4).

@@

Adult participants were recruited via Amazon Mechanical Turk (MTurk) and received payment equivalent to an hourly wage of ~ \$9. Each participant contributed data to only one experiment. Experiment 1 and each manipulation of experiment 2 had *N* = 40 participants. Sample size in experiment 3 was *N* = `r length(unique(adult_ex3_data$id))`. *N* = `r length(unique(adult_ex4_prior_data$id))` participated in the experiments to measure the strong, medium and weak preference and novelty manipulations that went into experiment 4. Finally, experiment 4 had *N* = `r length(unique(adult_ex4_data$id))` participants. Sample sizes in all adult experiments were chosen to yield at least 120 data points per cell. All studies were approved by the Stanford Institutional Review Board (protocol no. 19960).

@@

Experiments 1 and 2 confirmed that adults make pragmatic inferences based on information provided by the utterance as well as by common ground and provided quantitative estimates of the strength of these inferences for use in our model. We modeled the integration of utterance informativity and common ground as a process of socially-guided probabilistic inference, using the results of experiments 1 and 2 to inform key parameters of a computational model. The Rational Speech Act (RSA) model architecture introduced by Frank and Goodman [-@frank2012predicting] encodes conversational reasoning through the perspective of a listener (“he” pronoun) who is trying to decide on the intended meaning of the utterance he heard from the speaker (“she” pronoun). The basic idea is that the listener combines his uncertainty about the speaker’s intended meaning - a prior distribution over referents P(r) - with his generative model of how the utterance was produced: a speaker trying to convey information to him. To adapt this model to the word learning context, we enrich this basic architecture with a mechanism for expressing uncertainty about the meanings of words (lexical uncertainty) - a prior distribution over lexica P(L) [@bergen2016pragmatic].

@@

ex3_pragm_model_noise_param <- readRDS("../../stats/saves/ex3_pragm_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))%>%
  round(digits = 2)

ex3_flat_prior_model_noise_param <- readRDS("../../stats/saves/ex3_flat_prior_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))%>%
  round(digits = 2)

ex3_prior_only_model_noise_param <- readRDS("../../stats/saves/ex3_prior_only_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))%>%
  round(digits = 2)
```

In experiment 3, we combined the procedures of experiment 1 and 2A or 2B. The test setup was identical to experiment 1, however, before making a request, the speaker interacted with the objects so that some of them were preferred by or new to them (Fig \ref{fig:fig1}c). This combination resulted in two ways in which the two information sources could be aligned with one another. In the congruent condition, the object that was the more informative referent was also the one that was preferred by or new to the speaker. In the incongruent condition, the other object was the one that was preferred by or new to the speaker. Taken together, there were 2 (novelty or preference) x 2 (same or different speaker) x 2 (congruent or incongruent) = 8 conditions in experiment 3. For each of these eight conditions, we generated model predictions using the modelling framework introduced above. The test hypothesis about how information is integrated we compared the three models introduced in the introduction: The *integration model* in which both information sources are flexibly combined, the *no common ground model* that focused only on utterance-level information and the the *no informativeness model* that focused only on common ground information.

Participants completed eight trials for one of the common ground manipulations with two trials per condition (same/different speaker x congruent/incongruent). Conditions were presented in a random order. We discuss and visualize the results as the proportion with which participants chose the more informative object (i.e., the object that would be the more informative referent when only utterance information is considered).

@@

As a first step, we used a GLMM to test whether participants were sensitive to the different ways in which information could be aligned. We found that participants distinguished between congruent and incongruent trials when the speaker remained the same (model term: `alignment x speaker`; $\beta$ = `r lm_ex3_r%>%filter(term == "speakersame_speaker:alignmentincongruent")%>%pull(estimate)`, se = `r lm_ex3_r%>%filter(term == "speakersame_speaker:alignmentincongruent")%>%pull(std.error)`, *p* `r lm_ex3_r%>%filter(term == "speakersame_speaker:alignmentincongruent")%>%pull(p.value)`). Thus, participants were sensitive to the different combinations of manipulations.

@@

As a second step, we compared the model predictions to the data. Participants’ average responses were highly correlated with the  predictions from the *integration model* in each condition (Fig \ref{fig:fig3}b). When comparing model, we found that model fit was considerably better for the* integration model* compared to the *no common ground model* (Bayes Factor (BF) = `r format(ex3_bf%>%pull(pragmatic_noise_flat_prior_noise), digits = 2)`) or the *no informativeness model* (BF = `r format(ex3_bf%>%pull(pragmatic_noise_prior_only_noise), digits = 2)`), suggesting that participants considered and integrated both sources of information.

Finally, we examined the noise parameter for each model. The estimated proportion of random responses according to the *integration model* was `r ex3_pragm_model_noise_param%>%pull(mean)` (95% Highest Density Interval (HDI): `r ex3_pragm_model_noise_param%>%pull(ci_lower)` - `r ex3_pragm_model_noise_param%>%pull(ci_upper)`). This parameter was substantially lower for the *integration model* compared to the alternative models (*no common ground model*: `r ex3_flat_prior_model_noise_param%>%pull(mean)` [`r ex3_flat_prior_model_noise_param%>%pull(ci_lower)` - `r ex3_flat_prior_model_noise_param%>%pull(ci_upper)`]; *no informativeness model*: `r ex3_prior_only_model_noise_param%>%pull(mean)` [`r ex3_prior_only_model_noise_param%>%pull(ci_lower)` - `r ex3_prior_only_model_noise_param%>%pull(ci_upper)`]), lending additional support to the conclusion that the *integration model* better captured the behavioral data. Rather than explaining systematic structure in the data, the alternative models achieved their best fit only by assuming a very high level of noise.

@@

```{r model comparison ex4}
ex4_bf <- bind_rows(
  readRDS("../../stats/saves/ex4_prior_only_model_noise_loglike.rds") %>% mutate(model = "prior_only_noise"),
  readRDS("../../stats/saves/ex4_pragm_model_noise_loglike.rds") %>% mutate(model = "pragmatic_noise"),
  readRDS("../../stats/saves/ex4_flat_prior_model_noise_loglike.rds") %>% mutate(model = "flat_prior_noise")
)%>%
  group_by(model)%>%
  summarize(logP = logSumExp(value))%>%
  spread(model, logP) %>%
  mutate("pragmatic_noise_prior_only_noise" = exp(pragmatic_noise - prior_only_noise),
         "pragmatic_noise_flat_prior_noise" = exp(pragmatic_noise - flat_prior_noise),
         "prior_only_noise_flat_prior_noise" = exp(prior_only_noise - flat_prior_noise)) %>%
  select(-pragmatic_noise, -flat_prior_noise, -prior_only_noise)

ex4_pragm_model_noise_param <- readRDS("../../stats/saves/ex4_pragm_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))%>%
  round(digits = 2)

ex4_flat_prior_model_noise_param <- readRDS("../../stats/saves/ex4_flat_prior_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))%>%
  round(digits = 2)

ex4_prior_only_model_noise_param <- readRDS("../../stats/saves/ex4_prior_only_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))%>%
  round(digits = 2)

To test if the *integration model* makes accurate predictions for different combinations, we first replicated and then extended the results of experiment 3 to a broader range of experimental conditions. Specifically, we manipulated the strength of the common ground information (3 levels - strong, medium and weak - for preference and 2 levels - strong and medium - for novelty) by changing the way the speaker interacted with the objects prior to the request. The procedural details and statistical analysis for these these manipulations are described in the Supplementary Material available online. For experiment 4, we paired each level of prior strength manipulation with the informativeness inference in the same way as in experiment 3. This resulted in a total of 20 conditions, for which we generated a priori model predictions in the same way as in experiment 3. The strong prior manipulation in experiment 4 was a direct replication of experiment 3 (see Fig \ref{fig:fig3}c). Each participant was randomly assigned to a common ground manipulation and a level of prior strength and completed eight trials in total, two in each unique condition in that combination.

@@

The direct replication of experiment 3 within experiment 4 showed a very close correspondence between the two rounds of data collection (see Fig \ref{fig:fig3}c). GLMM results for experiment 4 can be found in the Supplementary Material available online. Here we focus on the analysis based on the probabilistic models. Model predictions from the *integration model* were again highly correlated with the average response in each condition (see Fig \ref{fig:fig3}d). We evaluated model fit for the same models as in experiment 3 and found again that the *integration model* fit the data much better compared to the *no common ground* (BF = `r format(ex4_bf%>%pull(pragmatic_noise_flat_prior_noise), digits = 2)`) or the *no informativeness model* (BF = `r format(ex4_bf%>%pull(pragmatic_noise_prior_only_noise), digits = 2)`). The inferred level of noise based on the data for the *integration model* was `r ex4_pragm_model_noise_param%>%pull(mean)` (95% HDI: `r ex4_pragm_model_noise_param%>%pull(ci_lower)` - `r ex4_pragm_model_noise_param%>%pull(ci_upper)`), which was similar to experiment 3 and again lower compared to the alternative models (*no common ground model*: `r ex4_flat_prior_model_noise_param%>%pull(mean)` [`r ex4_flat_prior_model_noise_param%>%pull(ci_lower)` - `r ex4_flat_prior_model_noise_param%>%pull(ci_upper)`]; *no informativeness model*: `r ex4_prior_only_model_noise_param%>%pull(mean)` [`r ex4_prior_only_model_noise_param%>%pull(ci_lower)` - `r ex4_prior_only_model_noise_param%>%pull(ci_upper)`]).

@@

The previous section showed that competent language users flexibly integrate information during pragmatic word learning. Do children make use of multiple information sources during word learning as well? When does this integration emerge developmentally? While many verbal theories of language learning imply that this integration takes place, the actual process has neither been described nor tested in detail. Here we provide an explanation in the form of our *integration model* and test if it is able to capture children’s word learning. Embedded in the assumptions of the model is the idea that developmental change is change in the strength of the individual inferences, leading to a change in the strength of the integrated inference. As a starting point, our model assumes developmental continuity in the integration process itself [@bohn2019pervasive], though this assumption could be called into question by a poor model fit. The study for children followed the same general pattern as the one for adults. We generated model predictions for how information should be integrated by first measuring children’s ability to use utterance (informativeness) and common ground (preference) information in isolation when making pragmatic inferences. We then adapted our model to study developmental change: We sampled children continuously between 3.0 and 5.0 years of age – a time in which children have been found to make the kind of pragmatic inferences we studied here [@bohn2019pervasive; @frank2014inferring] - and generated model predictions for the average developmental trajectory in each condition.

@@

In experiment 6, we assessed whether children use common ground information to identify the referent of a novel word. We tested children with the novelty as well as the preference manipulation. In both cases, we found that children made the basic inference in that they selected the object that was preferred by or new to the speaker, but found little evidence that children distinguished between requests made by the same speaker or a different speaker in the case of novelty. This finding contrasts with earlier work [diesendruck2004two]. However, since our focus was on how children selectively integrate informativeness and common ground, we did not follow up on this finding but dropped the novelty manipulation and focused on preference for the remainder of the study. The procedure for children was identical to the preference manipulation for adults. Children received eight trials, four with the same and four with a different speaker.

@@

In experiment 7, we combined the procedures of experiment 5 and 6 and collected new data from children between 3.0 and 5.0 years of age in each of the four conditions (Fig \ref{fig:fig1}c). We again inserted the preference manipulation into the setup of experiment 5. After greeting the child, the animal turned to one of the trees, pointed to an object (object was temporarily enlarged and moved closer to the animal) and expressed liking or disliking. Then the animal turned to the other tree and expressed the other attitude for the other kind of object. Next, the animal disappeared and either the same or a different animal returned. The rest of the trial was identical to the request phase of experiment 5. Children received eight trials, two per condition (same/different speaker x congruent/incongruent) in a randomized order.

@@

As a first step, we used a GLMM to test whether children were sensitive to the different ways in which information could be aligned. Children’s propensity to differentiate between congruent and incongruent trials for the same or a different speaker increased with age (model term: `age x alignment x speaker`; $\beta$ = `r child_lm_ex3_r%>%filter(term == "age_num:speakersame_speaker:alignmentincongruent")%>%pull(estimate)`, se = `r child_lm_ex3_r%>%filter(term == "age_num:speakersame_speaker:alignmentincongruent")%>%pull(std.error)`, *p* `r child_lm_ex3_r%>%filter(term == "age_num:speakersame_speaker:alignmentincongruent")%>%pull(p.value)`).

@@

child_ex3_pragm_model_noise_param <- readRDS("../../stats/saves/child_ex3_pragm_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

child_ex3_flat_prior_model_noise_param <- readRDS("../../stats/saves/child_ex3_flat_prior_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

child_ex3_prior_only_model_noise_param <- readRDS("../../stats/saves/child_ex3_prior_only_model_noise.rds") %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

Analyses comparing the model predictions from the probabilistic models to the data suggest that children flexibly integrate both common ground and informativity information. Furthermore, this integration process is accurately captured by the *integration model* at least for four-year-olds. For the correlational analysis, we binned model predictions and data by year. There was a substantial correlation between the predicted and measured average response for four-year-olds, but less so for three-year-olds (Fig \ref{fig:fig5}b). One of the reasons for the latter was the low variation between conditions. For the model comparison, we treated age continuously. As with adults, we found a much better model fit for the *integration model* compared to the *no common ground* (BF = `r format(child_ex3_bf%>%filter (parameter=="noise")%>%pull(pragmatic_flat_prior), digits = 2)`) or the *no informativeness model* (BF = `r format(child_ex3_bf%>%filter (parameter=="noise")%>%pull(pragmatic_prior_only), digits = 2)`).

The inferred level of noise based on the data for the integration model was `r child_ex3_pragm_model_noise_param%>%pull(mean)` (95% HDI: `r child_ex3_pragm_model_noise_param%>%pull(ci_lower)` - `r child_ex3_pragm_model_noise_param%>%pull(ci_upper)`), which was lower compared to the alternative models considered (*no common ground model*: `r child_ex3_flat_prior_model_noise_param%>%pull(mean)` [`r child_ex3_flat_prior_model_noise_param%>%pull(ci_lower)` - `r child_ex3_flat_prior_model_noise_param%>%pull(ci_upper)`]; *no informativeness model*: `r child_ex3_prior_only_model_noise_param%>%pull(mean)` [`r child_ex3_prior_only_model_noise_param%>%pull(ci_lower)` - `r child_ex3_prior_only_model_noise_param%>%pull(ci_upper)`]) but numerically higher than that of adults.

The high level of inferred noise moved the model predictions for children in all conditions close to chance level. We therefore compared two additional sets of models with different parameterizations of the noise parameter that emphasized differences between conditions in the model predictions more (see Supplementary Material available online and Fig \ref{fig:fig5}a). This analysis was not pre-registered. Parameter free models did not include a noise parameter and developmental noise models allowed the noise parameter to change with age. In each case, the *integration model* provided a better fit compared to the alternative models (*no common ground*: parameter free BF = `r format(child_ex3_bf%>%filter (parameter=="parameter free")%>%pull(pragmatic_flat_prior), digits = 2)`, developmental noise BF = `r format(child_ex3_bf%>%filter (parameter=="developmental noise")%>%pull(pragmatic_flat_prior), digits = 2)`; *no informativeness*: parameter free BF = `r format(child_ex3_bf%>%filter (parameter=="parameter free")%>%pull(pragmatic_prior_only), digits = 2)`, developmental noise BF = `r format(child_ex3_bf%>%filter (parameter=="developmental noise")%>%pull(pragmatic_prior_only), digits = 2)`). The developmental noise parameter for the integration model decreased with age, suggesting that older children behaved more in line with model predictions compared to younger children (see Fig. S13 in Supplementary Material available online).

@@

Integrating multiple sources of information is an integral part of human communication [@tanenhaus1995integration]. To infer the intended meaning of an utterance, listeners must combine their knowledge of communicative conventions (semantics and syntax) with social expectations about their interlocutor. This integration is especially vital in early language learning, and the different varieties of pragmatic information are among the most important sources [@bohn2019pervasive]. But how are different cues integrated during word learning? Here we used a Bayesian cognitive model to formalize this integration process. We studied how utterance-level (Gricean) expectations about informative communication are integrated with common ground information. Adults’ and children’s learning was best predicted by a model in which both sources of information traded-off flexibly. Alternative models that considered only one source of information made substantially worse predictions.

All of the models we compared here integrated some explicit structure, rather than (for example) simply weighing information sources by some ratio. We made this decision because we wanted to make predictions within a framework in which the models were models of the task, rather than simply models of the data. That is, inferences are not computed separately by the modeler and specified as inputs to a regression model, but instead are the results of an integrated process that operates over a (schematic) representation of the experimental stimuli. Further, our models are variants derived from the broader RSA framework, which has been integrated into larger systems for language learning in context [@wang2016learning].

How is information integrated in this context in this context? The *integration model* assumes that the informativeness of an utterance depends on the common ground shared between interlocutors. That is, the listener assumes that the speaker takes the common ground shared between the speaker and the (naive) listener as a starting point when computing the effect of each utterance. As a consequence, when prior interactions strongly implicate one object as the more likely referent (for example in the preference - same speaker conditions in experiment 3, 4 and 7), the speaker reasons that this object will be the inferred referent of any semantically plausible utterance, even when the same utterance would point to a different object in the absence of common ground. Taken together, this model provides an explicit and formal description of the integration process, thereby offering an answer to the question of *how* information may be integrated during pragmatic word learning. Predictions generated based on this process accurately captured adults' inferences across a wide range of conditions.

The *integration model* also predicted information integration in four-year-olds. However, the model did not successfully describe three-year-olds’ inferences; thus, it is possible that they were not able to integrate information sources. But our findings are also consistent with a simpler explanation, namely that the overall weaker responses we observed in the independent measurement experiments (experiments 5 and 6), combined with some noise in responding, led the younger children to appear relatively random in their responses. It would therefore be important to follow up with a combination of tasks that produce stronger inferences in younger children.

The primary source of developmental change in our model are age related changes in the propensity to make the individual inferences. As they get older, children expect speakers to be more informative and to be more likely to follow common ground, but the process by which the two information sources are integrated at any given age is assumed to be the same. The developmental noise model reported for experiment 7 offers a complimentary way to think about developmental change. We found that the estimated level of noise for the *integration model* dropped with age. That is, while, according to this model, young children's responses were best explained by random guessing, older children responded in line with the model structure. One way to interpret this would be to say that handling and integrating the two information sources poses a greater challenge to younger children. As soon as the necessary abilities are in place, children flexibly combine the two information sources as described by the *integration model*. This points to domain general processes as an additional locus of developmental but also assumes that the integration process itself is stable across age. Other developmental models are also worth exploring in future work; one possible candidate would be a model in which the integration process itself changes with age.

@@

# Declarations of interest

None.
